Notes on Validated Model Counting

### 2021-04-05 doc/mc-notes.tex
Improved theorem presentation
Minor edits

\begin{lemma}
\label{lemma:independent:split}
if Boolean formula $F$ is independent of variable $x$, then:
\begin{equation}
|\interpset{F}| \; = \; 2\cdot |\interpset{F\land x}| = \; 2\cdot |\interpset{F\land \obar{x}}| \label{eqn:independent}
\end{equation}  
\end{lemma}  
That is, the set of interpretations of $F$ are split evenly between
those assigning $\tautology$ to $x$ and those assigning $\nil$ to $x$.

@@

Letting $n = |X|$, we can use the following recursive computation:
\begin{eqnarray}
\mcount(\tautology) & = & 2^{n} \label{eqn:count:true}\\
\mcount(\nil) & = & 0 \label{eqn:count:false}\\
\mcount(x) & = & 2^{n-1} \label{eqn:count:variable}\\
\mcount(\ite(x, v_t, v_e)) & = & \frac{\mcount(v_t) + \mcount(v_e)}{2} \label{eqn:count:ite}
\end{eqnarray}

\begin{theorem}
Equations (\ref{eqn:count:true}--\ref{eqn:count:ite}) correctly hold for any node $v$ in a free ITE graph.
\end{theorem}  
\noindent
{\em Proof:} The key result here is (\ref{eqn:count:ite}).  It follows
from Lemma~\ref{lemma:independent:split}, (\ref{eqn:ite:interp}),
Lemma~\ref{lemma:independent:dset}, and (\ref{eqn:depend:ite}).  That is, both
$v_t$ and $v_e$ are independent of $x$.  That implies that half of the
satisfying assignments $\interp$ for $v_t$ have have $\interp(x) =
\tautology$.  Similarly, half of the satisfying assignments $\interp$
for $v_e$ have $\interp(x) = \nil$.  Equation (\ref{eqn:count:ite}) then follows by (\ref{eqn:ite:interp}).

>>>>

\begin{lemma}
\label{lemma:independent:split}
If Boolean formula $F$ is independent of variable $x$, then:
\begin{equation}
|\interpset{F}| \; = \; 2\cdot |\interpset{F\land x}| = \; 2\cdot |\interpset{F\land \obar{x}}| \label{eqn:independent}
\end{equation}  
\end{lemma}  
That is, the set of models of $F$ is split evenly between
those assigning $\tautology$ to $x$ and those assigning $\nil$ to $x$.

@@

\begin{theorem}
\label{theorem:count}
Letting $n = |X|$, the following recursive formula holds for any node in a free ITE graph
\begin{eqnarray}
\mcount(\tautology) & = & 2^{n} \label{eqn:count:true}\\
\mcount(\nil) & = & 0 \label{eqn:count:false}\\
\mcount(x) & = & 2^{n-1} \label{eqn:count:variable}\\
\mcount(\ite(x, v_t, v_e)) & = & \frac{\mcount(v_t) + \mcount(v_e)}{2} \label{eqn:count:ite}
\end{eqnarray}  
\end{theorem}
\noindent
{\em Proof:} The key result here is (\ref{eqn:count:ite}).  It follows
from Lemma~\ref{lemma:independent:split}, (\ref{eqn:ite:interp}),
Lemma~\ref{lemma:independent:dset}, and (\ref{eqn:depend:ite}).  That is, both
$v_t$ and $v_e$ are independent of $x$.  That implies that half of the
satisfying assignments $\interp$ for $v_t$ have have $\interp(x) =
\tautology$.  Similarly, half of the satisfying assignments $\interp$
for $v_e$ have $\interp(x) = \nil$.  Equation (\ref{eqn:count:ite}) then follows by (\ref{eqn:ite:interp}).
